Data Management \\ and LSST Special Programs

### 2017-11-14 DMTN-065.tex
Cite documents in standard manner

The intent of the LSST Data Management team for processing data from Special Programs can be summarized by the following statement, which reflects Section 6 of the Data Products Definitions Document (DPDD, LSE-163, \cite{LSE-163}; but see also the minor proposed change to this statement in \ref{Size-1}):

@@

The current technical limitations on what kind of observations can be obtained with the LSST telescope and camera are described below in a rough draft, to be updated in the future. Additionally, in Section \ref{ssec:docrev_oss}, we review the Observatory System Specifications (OSS, LSE-30, \cite{LSE-30}) document and highlight a couple more questions regarding potential boundaries on Special Programs observations.

Filter Changes: \\
$\bullet$ The maximum time for filter change is 120 seconds: 30 seconds for the telescope to reorient the camera to its nominal zero angle position on the rotator, and 90 seconds to the camera subsystem for executing the change (OSS-REQ-0293). \\
$\bullet$ Note that changes regarding the requirements on filters in support of "deep drilling fields" are underway here: \url{https://project.lsst.org/groups/ccb/node/938}.
$\bullet$ \textit{(MLG - The next three points are from a document-in-progress by Ivezic and Ritz; update to formal citation when it's public-facing.)} \\
$\bullet$ The minimum time between filter changes has no restrictions from e.g., thermal tolerances. However, based on overheads and efficiency, it is recommended to keep the filter change rate lower than once every 10 minutes. \\
$\bullet$ The maximum total number of filter changes is 100,000 over 15 years, an average of 18 changes per night. \\
$\bullet$ The maximum number of filter swaps in/out of the carousel is 3000 in 15 years, or once every two nights. \\

@@

As presented in Section \ref{sec:intro}, the intent of LSST DM for processing data from Special Programs is that ``LSST will not write unique algorithms for processing Special Programs data or reprocessing Main Survey data, but LSST will reconfigure the pipelines and generate imaging and catalog products for Special Programs data whenever possible" (DPDD, LSE-163, \cite{LSE-163}), and that an additional $\sim$10\% of the LSST computational system has been added and will be dedicated to user processing (of main survey or special programs data).

@@

A standard visit image is considered to be two snap exposures of $\sim15$ second each, and in the WFD main survey sequential visits will be of different fields. As we found in Section \ref{sec:data}, Special Programs are most likely to request observing modes with shorter exposure times and long sequences of visits to the same field. In this section we review DM's planned capability to process such non-standard visit sequences, keeping in mind that DM's processing boundaries might ultimately be defined not by exposure acquisition parameters, but by the image quality parameters such as number of stars with sufficient flux for calibration or the source area density (crowding), and that these boundaries might not be fully constrained until the final performance of DM algorithms as described in the Data Management Applications Design (DMAD, LDM-151, \cite{LDM-151}) document are characterized (see also Section \ref{ssec:docrev_dmad}).

@@

In a ``possible" scenario, DM would assemble a pipeline from existing DM codes in order to process data associated with a given Special Program and build image and catalog products that meet the science needs of that particular program. For example, for a DDF SN survey (see Section \ref{ssec:SPCS_SNDDF}), existing DM codes would be used to: (1) make a deep template image from a certain time window, (2) process standard single visit images, (3) create a nightly CoAdd, (4) run difference imaging analysis, (5) run source detection on the difference images, and (6) create \texttt{DIASource} and \texttt{DIAObject} catalog equivalents (this example is also given in Section 6 of the DPDD, \cite{LSE-163}). This type of reconfiguration is also be possible for users to do as a Level 3 pipeline (see below), but having these products provided by DM ensures a consistent and verified level of quality and public access to all processed Special Programs data products, and is expected to be a large benefit to the science results.

@@

% % % % % % %
\subsection{Science Requirements Document, SRD, LPM-17, \cite{LPM-17}}\label{ssec:docrev_srd}

@@

% % % % % % %
\subsection{LSST System Requirements, LSR, LSE-29, \cite{LSE-29}}\label{ssec:docrev_lsr}

@@

% % % % % % %
\subsection{Observatory System Specifications, OSS, LSE-30, \cite{LSE-30}}\label{ssec:docrev_oss}

@@

% % % % % % %
\subsection{Data Management Subsystems Requirements, DMSR, LSE-61, \cite{LSE-61}}\label{ssec:docrev_dmsr}

@@

% % % % % % %
\subsection{Data Management Applications Design, DMAD, LDM-151, \cite{LDM-151}}\label{ssec:docrev_dmad}

@@

% % % % % % %
\subsection{Data Products Definitions Document, DPDD, LSE-163, \cite{LSE-163}}\label{ssec:docrev_dpdd}

@@

% % % % % % %
\subsection{Photometric Calibration for the LSST Survey, LSE-180, \cite{LSE-180}}

@@

So far, only one aspect of the LSST Special Programs are set: the locations of the four chosen deep drilling fields\footnote{\url{https://www.lsst.org/scientists/survey-design/ddf}}. \cite{2008arXiv0805.2366I} describes a nominal DDF data set as e.g., a series of fifty 15 second exposures in each of the $griz$ filters obtained once every two nights for four months, which would yield a nightly stack depth of $r<26.5$ and a full stack depth of $r<28.0$ (once combined with the wide-fast-deep survey data). Additionally, three mini-survey areas that have already been considered in detail are likely to move forward: the North Ecliptic Spur (NES), the South Celestial Pole, and the Galactic Plane (see Figure 8 of \cite{2008arXiv0805.2366I}). We have a summary collection of past white papers and informally proposed mini-surveys in Section \ref{ssec:data_science}.

@@

\noindent \textbf{Resources:} \\
$\bullet$ "LSST: from Science Drivers to Reference Design and Anticipated Data Products" Ivezi\'{c} et al. (2008), \cite{2008arXiv0805.2366I} \\
$\bullet$ The LSST Science Requirements Document (SRD), LPM-17, \cite{LPM-17} \\
$\bullet$ LSST Deep Drilling white papers from 2011: \url{https://project.lsst.org/content/whitepapers32012} \\
$\bullet$ "General Review of the Proposed DDF and MS", LSST AHM Aug 2016 presentation by Niel Brandt \url{https://project.lsst.org/meetings/lsst2016/sites/lsst.org.meetings.lsst2016/files/Brandt-DDF-MiniSurveys-01.pdf} \\
$\bullet$ "Simulations, Metrics and Merit Function for Mini-Surveys and DDF", LSST AHM Aug 2016 presentation by Stephen Ridgway \url{https://project.lsst.org/meetings/lsst2016/sites/lsst.org.meetings.lsst2016/files/Ridgway-SimulationsMetrics_1.pdf} \\
$\bullet$ "LSST's DC [Deep CoAdd] Bias Against Planets and Galactic-Plane Science" by A. Gould, \cite{2013arXiv1304.3455G} \url{https://arxiv.org/abs/1304.3455} \\
$\bullet$ Chapter 10 "Special Surveys" of the Observing Strategy White Paper \cite{2017arXiv170804058L}

@@

\textbf{A Nominal DDF Observing Strategy -- } Ivezi\'{c} et al. (2008, \cite{2008arXiv0805.2366I}), Section 3.2.1 "Mini-Surveys": describes a nominal DDF data set as $\sim50$ consecutive $15$ second exposures in each of four filters in one hour per night, once every two nights, for four months. Each observation would have a limit of $r\sim24.5$; a one-hour nightly stack would have a limit of $r\sim26.5$; and and assuming a $60\%$ completion rate (weather), the four-month $\sim40$ hours stacked together with the $\sim180$ main survey visits would yield a limit of $r\sim28$. (Note that this nominal survey is missing the overhead for readout and shutter, 3 seconds per exposure, which means that the LSST capability is closer to 43--45 exposures per hour, minus the time for 3--4 filter changes -- not 50 exposures per hour.)

\medskip
\noindent \textbf{Solar System Objects (SSO)}\\
$\bullet$ North Ecliptic Spur: observations in this area of the sky will yield more $\geq140$ m near-earth objects (NEOs) for the final LSST sample (Brandt talk). \\
$\bullet$ Mini-Moons: finding and studying temporarily captured satellites of the Earth. (Section 10.2, \cite{2017arXiv170804058L}) \\
$\bullet$ Sweetspot: use twilight fields to find NEOs in Earth-like orbits (never in opposition fields). (Section 10.2, \cite{2017arXiv170804058L}) \\
$\bullet$ Meter-Sized Impactors: finding and tracking meter-sized impactors $<2$ weeks before impact.  (Section 10.2, \cite{2017arXiv170804058L}) \\
$\bullet$ Detecting faint SSO down to $r\sim27$ would be done by applying shift-and-stack (SAS) processing to a mini-survey data of standard visit images (faint SSO include: Trans-Neptunian Objects Trojans, asteroids, long-period comets, dwarf planets) \citep{BeckerWP}. Regarding SAS, \cite{BeckerWP} says that "the multi-fit algorithm ... naturally provides a base infrastructure for this process. In particular, the marshaling of the pixels to attempt a given photometric measurement is non-trivial when tens of thousands of images are required. However, the multi-fit middleware is required to do exactly this, so we expect that this issue will be resolved by the time SAS is needed." \\
$\bullet \bullet \bullet$ \textbf{Summary.} Most of these science goals can be achieved with standard visit images obtained in a normal survey pattern taken during the night. The exception is a Sweetspot survey, which occurs during twilight and thus may require shorter exposures. Many of these science goals will also be possible by using the products of Moving Object Processing System (MOPS), which runs on the Level 1 \texttt{DIASource} catalogs updated each night. The exception is finding faint SSOs through shift-and-stack processing; SAS is not a capability of the DM system, cannot be done solely by reconfiguring DM pipelines, and is therefore considered Level 3. See Sections \ref{ssec:SPCS_SAS} and \ref{ssec:SPCS_TNO} for more detailed DM processing case studies.

\noindent \textbf{Stars in the Milky Way and Magellanic Clouds} \\
$\bullet$ When surveying the Galactic Plane, obtaining all the images over a short time span would mitigate proper motion loss (and increase flare detection rates) and help identify useful stellar populations. When targeting Galactic plane regions and/or open clusters, typically 1 to 3 filters are needed for detections of e.g., faint stars already detected in $z$ and $y$ but needing $i$ to distinguish from red galaxies \cite{DhitalWP}. \\
$\bullet$ A nominal deep-drilling observing strategy over the full L/SMC galaxies will characterize stellar variability to $M_V<6.5$ on timescales from 15s to 3d \cite{SzkodyWP}. For this, special co-adds may be required, e.g., \textit{"to reach variability levels of 0.1 to 0.005 mag will require co-adds depending on the timescale of the particular variables"}. \citep{SzkodyWP}. \\
$\bullet$ A Twilight Short Exposure survey would enable bright stars, which have longer monitoring baselines due to historical observations, to be put on the same photometric system as the deeper LSST WFD main survey catalog. \\
$\bullet$ A short-exposure survey of M67, Chapter 10.4 of \cite{2017arXiv170804058L}, suggests using the stretch goal of 0.1 second exposures or, if that is not possible, \textit{"custom pixel masks to accurately perform photometry on stars as much as 6 magnitudes brighter than the saturation level"}. \\
$\bullet \bullet \bullet$ \textbf{Summary.} Many of these science goals can be achieved with standard visit images, but with alternative observing strategies such as limiting the number of filters or concentrating the depth build-up into shorter timeframes; the exception is short-exposure surveys to obtain bright stars. Similarly, most of these science goals can be met by reconfiguring the planned DM pipelines to detect sources, measure proper motion, and create deep stacks. The exception might be in processing exposures as short as 0.1 seconds, and attempting to measure flux of a saturated stars. See Section \ref{ssec:SPCS_GPVSEx} for more detailed DM processing case studies.

\noindent \textbf{Exoplanets} \\
$\bullet$ Transits. The nominal DDF plan described in \cite{2008arXiv0805.2366I} would allow for $1\%$ variability detection over hour-long timescales, which is suitable for detecting transits. A DDF field at Galactic latitude $30$ degrees would yield $10^6$ stars at $r<21$ that would have $\mathrm{SNR}>100$ in each single exposure of the sequence (as described in Section 3.1.2 of \cite{2008arXiv0805.2366I}). \cite{2013arXiv1304.3455G} describes how transits can be extract from a wider-area survey of the Galactic Plane.\\
$\bullet$ Microlensing.  Slower than a transit, \cite{2013arXiv1304.3455G} suggests that $\sim22$ mag imaging over the Galactic Plane region every 3-4 days (i.e., the WFD nominal cadence) can find microlensing candidates, which would then need follow-up with external facilities. However, this will require image differencing in crowded fields. \cite{2013arXiv1304.3455G} suggests that the Galactic Plane can yield a lot of science despite the fact that its eventual deep co-adds would be uselessly confusion limited, and therefore should not be skipped. Microlensing can also be done with DDF-style observing at low Galactic latitudes. \\
$\bullet \bullet \bullet$ \textbf{Summary.} All of these science goals appear possible with standard visit images. They might be possible with reconfigured DM pipelines, but this depends heavily on performance in crowded fields. See Section \ref{ssec:SPCS_GPVSEx} for a more detailed DM processing case study for Galactic Plane regions.

\noindent \textbf{Supernovae} \\
$\bullet$ The nominal DDF plan described in \cite{2008arXiv0805.2366I}, which builds nightly stacks with a limit of $r\sim26.5$ out of standard visit images, would extend the SN sample to $z\sim1.2$ and provide more densely sampled light curves for cosmological analyses. The optimal exposure time distribution might be 6, 5, 10, 10, 9, 10 in $ugrizy$. \cite{KesslerWP} \\
$\bullet$ High-cadence observations of DDF would be the only way to detect fast transients, particularly extragalactic novae, some tidal disruption events, optical counterparts to gamma-ray bursts, and peculiar SNe \cite{2014ApJ...794...23D}. \\
$\bullet$ Generating the best-possible individual SN light curves for cosmological analyses requires building special, deep-as-possible, SN-free host galaxy images and using them as a template. This will also be necessary for studying SNe that appear in the template image; i.e., that last $>1000$ days. These are mostly Type IIn, probably explosions of massive stars into dense circumstellar material, which are not used for cosmology but rather to study late-stage stellar evolution and mass loss. SN-free images will also be needed to measure correlated properties for cosmology and to do host-galaxy science. The latter, specifically the "characterization of ultra-faint SN host galaxies", is also mentioned in the Galaxies DDF WP \cite{FergusonWP}. \\
$\bullet$ \textit{MLG - I'm not sure this has been yet proposed, but short-exposure observations to include nearby SNe on the same photometric system could be useful?} \\
$\bullet \bullet \bullet$ \textbf{Summary.} All of these science goals appear possible with standard visit images (with the exception of an as-yet unproposed target-of-opportunity survey to obtain photometry for nearby, saturated SNe). Additionally, all are accessible with reconfigured DM pipelines to stack and difference the data. In particular, the Level 2 DRP codes to create "transient-free CoAdds" will be suitable for generating the SN-free templates for DDF, as they will do for the Main Survey images. See also Section \ref{ssec:SPCS_SNDDF} for a DM processing case study to find SNe in a DDF.

\noindent \textbf{Galaxies} \\
$\bullet$ Use the additional depth of a DDF to build a large collection of low-$\mu$ objects. \cite{FergusonWP} mentions "identification of nearby isolated low-redshift dwarf galaxies via surface-brightness fluctuations" and "characterization of low-surface-brightness extended features around both nearby and distant galaxies". \\
$\bullet$ The DDF stacks can also be used to characterize of high-$z$ clusters, although this ability might depend on deblending extended objects. \\
$\bullet$ The DDF observations, when combined with the WFD, allow for AGN monitoring on a variety of timescales in well-characterized galaxies \citep{FergusonWP,GawiserWP}. \\
$\bullet \bullet \bullet$ \textbf{Summary.} As with the SN science goals, these use standard visit images and reconfigured DM pipelines to make deep CoAdds and extract sources. In addition, it seems likely that Level 3 algorithms that are optimized to detect and characterize particular types of faint extended sources will be needed, but these are beyond the scope of DM.

@@

The document "Cadence Exploration and Improvement Plan" (by Zeljko, Lynne, Andy) proposes that the next call for DDF White Papers (proposed fields and cadences) be in December 2017, and the call for Mini-Survey White Papers be in October 2018. At the 2017 All Hands Meeting the Science Advisory Council seemed likely to recommend these be combined into a single call. The relevant Jira issue is \url{https://jira.lsstcorp.org/browse/LIT-325?filter=-3} and the Project's initial thoughts regarding this call for white papers are described by Beth Willman at the LSST SAC Meeting in October 2016\footnote{\url{https://project.lsst.org/groups/sac/sites/lsst.org.groups.sac/files/Willman_DDF.pdf}}. The plan for ingesting these white paper proposals is for them to be reviewed by the Science Advisory Council based on a set of criteria set by the Project Science Team, and the SAC will make recommendations to the LSST Director. The criteria to be met by the proposals may include specifications that they: satisfy minimum technical requirements (feasibility, overheads); maximize diverse scientific objectives (serve a wide community); and generate legacy datasets and add value to the products of other facilities. After the proposals have been reviewed (and perhaps culled or combined, as recommended by the SAC) it is projected that new OpSim runs will be produced in mid-2019 for further evaluation. While the initial white paper proposals will be separate from the existing Observing Strategy White Paper \cite{2017arXiv170804058L}, they could be incorporated into a comprehensive analysis of the options when these simulations are ready.

@@

A short-exposure survey of the bright stars of M67, described in Chapter 10.4 of the Observing Strategy White Paper \cite{2017arXiv170804058L}, suggests using the stretch goal of 0.1 second exposures or, if that is not possible, \textit{"custom pixel masks to accurately perform photometry on stars as much as 6 magnitudes brighter than the saturation level"}. This would certainly be considered a Level 3 algorithm.

>>>>

The intent of the LSST Data Management team for processing data from Special Programs can be summarized by the following statement, which reflects Section 6 of the Data Products Definitions Document (\DPDD, \citeds{LSE-163}; but see also the minor proposed change to this statement in \ref{Size-1}):

@@

The current technical limitations on what kind of observations can be obtained with the LSST telescope and camera are described below in a rough draft, to be updated in the future. Additionally, in Section \ref{ssec:docrev_oss}, we review the Observatory System Specifications (OSS, \citeds{LSE-30}) document and highlight a couple more questions regarding potential boundaries on Special Programs observations.

Filter Changes: \\
$\bullet$ The maximum time for filter change is 120 seconds: 30 seconds for the telescope to reorient the camera to its nominal zero angle position on the rotator, and 90 seconds to the camera subsystem for executing the change (OSS-REQ-0293). \\
$\bullet$ Note that changes regarding the requirements on filters in support of "deep drilling fields" are underway here: \href{https://project.lsst.org/groups/ccb/node/938}{LCR-458}. \\
$\bullet$ \textit{(MLG - The next three points are from a document-in-progress by Ivezic and Ritz; update to formal citation when it's public-facing.)} \\
$\bullet$ The minimum time between filter changes has no restrictions from e.g., thermal tolerances. However, based on overheads and efficiency, it is recommended to keep the filter change rate lower than once every 10 minutes. \\
$\bullet$ The maximum total number of filter changes is 100,000 over 15 years, an average of 18 changes per night. \\
$\bullet$ The maximum number of filter swaps in/out of the carousel is 3000 in 15 years, or once every two nights. \\

@@

As presented in Section \ref{sec:intro}, the intent of LSST DM for processing data from Special Programs is that ``LSST will not write unique algorithms for processing Special Programs data or reprocessing Main Survey data, but LSST will reconfigure the pipelines and generate imaging and catalog products for Special Programs data whenever possible" (\DPDD, \citeds{LSE-163}), and that an additional $\sim$10\% of the LSST computational system has been added and will be dedicated to user processing (of main survey or special programs data).

@@

A standard visit image is considered to be two snap exposures of $\sim15$ second each, and in the WFD main survey sequential visits will be of different fields. As we found in Section \ref{sec:data}, Special Programs are most likely to request observing modes with shorter exposure times and long sequences of visits to the same field. In this section we review DM's planned capability to process such non-standard visit sequences, keeping in mind that DM's processing boundaries might ultimately be defined not by exposure acquisition parameters, but by the image quality parameters such as number of stars with sufficient flux for calibration or the source area density (crowding), and that these boundaries might not be fully constrained until the final performance of DM algorithms as described in the Data Management Applications Design (DMAD, \citeds{LDM-151}) document are characterized (see also Section \ref{ssec:docrev_dmad}).

@@

In a ``possible" scenario, DM would assemble a pipeline from existing DM codes in order to process data associated with a given Special Program and build image and catalog products that meet the science needs of that particular program. For example, for a DDF SN survey (see Section \ref{ssec:SPCS_SNDDF}), existing DM codes would be used to: (1) make a deep template image from a certain time window, (2) process standard single visit images, (3) create a nightly CoAdd, (4) run difference imaging analysis, (5) run source detection on the difference images, and (6) create \texttt{DIASource} and \texttt{DIAObject} catalog equivalents (this example is also given in Section 6 of the \DPDD, \citedsp{LSE-163}). This type of reconfiguration is also be possible for users to do as a Level 3 pipeline (see below), but having these products provided by DM ensures a consistent and verified level of quality and public access to all processed Special Programs data products, and is expected to be a large benefit to the science results.

@@

% % % % % % %
\subsection{Science Requirements Document, SRD, \citeds{LPM-17}}\label{ssec:docrev_srd}

@@

% % % % % % %
\subsection{LSST System Requirements, LSR, \citeds{LSE-29}}\label{ssec:docrev_lsr}

@@

% % % % % % %
\subsection{Observatory System Specifications, OSS, \citeds{LSE-30}}\label{ssec:docrev_oss}

@@

% % % % % % %
\subsection{Data Management Subsystems Requirements, DMSR, \citeds{LSE-61}}\label{ssec:docrev_dmsr}

@@

% % % % % % %
\subsection{Data Management Applications Design, DMAD, \citeds{LDM-151}}\label{ssec:docrev_dmad}

@@

% % % % % % %
\subsection{Data Products Definitions Document, DPDD, \citeds{LSE-163}}\label{ssec:docrev_dpdd}

@@

% % % % % % %
\subsection{Photometric Calibration for the LSST Survey, \citeds{LSE-180}}

@@

So far, only one aspect of the LSST Special Programs are set: the locations of the four chosen deep drilling fields\footnote{\url{https://www.lsst.org/scientists/survey-design/ddf}}. \citep{2008arXiv0805.2366I} describes a nominal DDF data set as e.g., a series of fifty 15 second exposures in each of the $griz$ filters obtained once every two nights for four months, which would yield a nightly stack depth of $r<26.5$ and a full stack depth of $r<28.0$ (once combined with the wide-fast-deep survey data). Additionally, three mini-survey areas that have already been considered in detail are likely to move forward: the North Ecliptic Spur (NES), the South Celestial Pole, and the Galactic Plane (see Figure 8 of \citep{2008arXiv0805.2366I}). We have a summary collection of past white papers and informally proposed mini-surveys in Section \ref{ssec:data_science}.

@@

\noindent \textbf{Resources:} \\
$\bullet$ "LSST: from Science Drivers to Reference Design and Anticipated Data Products" Ivezi\'{c} et al. (2008), \citep{2008arXiv0805.2366I} \\
$\bullet$ The LSST Science Requirements Document (\SRD), \citeds{LPM-17} \\
$\bullet$ LSST Deep Drilling white papers from 2011: \url{https://project.lsst.org/content/whitepapers32012} \\
$\bullet$ "General Review of the Proposed DDF and MS", LSST AHM Aug 2016 presentation by Niel Brandt \url{https://project.lsst.org/meetings/lsst2016/sites/lsst.org.meetings.lsst2016/files/Brandt-DDF-MiniSurveys-01.pdf} \\
$\bullet$ "Simulations, Metrics and Merit Function for Mini-Surveys and DDF", LSST AHM Aug 2016 presentation by Stephen Ridgway \url{https://project.lsst.org/meetings/lsst2016/sites/lsst.org.meetings.lsst2016/files/Ridgway-SimulationsMetrics_1.pdf} \\
$\bullet$ "LSST's DC [Deep CoAdd] Bias Against Planets and Galactic-Plane Science" by A. Gould, \citep{2013arXiv1304.3455G} \url{https://arxiv.org/abs/1304.3455} \\
$\bullet$ Chapter 10 "Special Surveys" of the Observing Strategy White Paper \citep{2017arXiv170804058L}

@@

\textbf{A Nominal DDF Observing Strategy -- } Ivezi\'{c} et al. (2008, \citep{2008arXiv0805.2366I}), Section 3.2.1 "Mini-Surveys": describes a nominal DDF data set as $\sim50$ consecutive $15$ second exposures in each of four filters in one hour per night, once every two nights, for four months. Each observation would have a limit of $r\sim24.5$; a one-hour nightly stack would have a limit of $r\sim26.5$; and and assuming a $60\%$ completion rate (weather), the four-month $\sim40$ hours stacked together with the $\sim180$ main survey visits would yield a limit of $r\sim28$. (Note that this nominal survey is missing the overhead for readout and shutter, 3 seconds per exposure, which means that the LSST capability is closer to 43--45 exposures per hour, minus the time for 3--4 filter changes -- not 50 exposures per hour.)

\medskip
\noindent \textbf{Solar System Objects (SSO)}\\
$\bullet$ North Ecliptic Spur: observations in this area of the sky will yield more $\geq140$ m near-earth objects (NEOs) for the final LSST sample (Brandt talk). \\
$\bullet$ Mini-Moons: finding and studying temporarily captured satellites of the Earth. (Section 10.2, \citep{2017arXiv170804058L}) \\
$\bullet$ Sweetspot: use twilight fields to find NEOs in Earth-like orbits (never in opposition fields). (Section 10.2, \citep{2017arXiv170804058L}) \\
$\bullet$ Meter-Sized Impactors: finding and tracking meter-sized impactors $<2$ weeks before impact.  (Section 10.2, \citep{2017arXiv170804058L}) \\
$\bullet$ Detecting faint SSO down to $r\sim27$ would be done by applying shift-and-stack (SAS) processing to a mini-survey data of standard visit images (faint SSO include: Trans-Neptunian Objects Trojans, asteroids, long-period comets, dwarf planets) \citep{BeckerWP}. Regarding SAS, \citep{BeckerWP} says that "the multi-fit algorithm ... naturally provides a base infrastructure for this process. In particular, the marshaling of the pixels to attempt a given photometric measurement is non-trivial when tens of thousands of images are required. However, the multi-fit middleware is required to do exactly this, so we expect that this issue will be resolved by the time SAS is needed." \\
$\bullet \bullet \bullet$ \textbf{Summary.} Most of these science goals can be achieved with standard visit images obtained in a normal survey pattern taken during the night. The exception is a Sweetspot survey, which occurs during twilight and thus may require shorter exposures. Many of these science goals will also be possible by using the products of Moving Object Processing System (MOPS), which runs on the Level 1 \texttt{DIASource} catalogs updated each night. The exception is finding faint SSOs through shift-and-stack processing; SAS is not a capability of the DM system, cannot be done solely by reconfiguring DM pipelines, and is therefore considered Level 3. See Sections \ref{ssec:SPCS_SAS} and \ref{ssec:SPCS_TNO} for more detailed DM processing case studies.

\noindent \textbf{Stars in the Milky Way and Magellanic Clouds} \\
$\bullet$ When surveying the Galactic Plane, obtaining all the images over a short time span would mitigate proper motion loss (and increase flare detection rates) and help identify useful stellar populations. When targeting Galactic plane regions and/or open clusters, typically 1 to 3 filters are needed for detections of e.g., faint stars already detected in $z$ and $y$ but needing $i$ to distinguish from red galaxies \citep{DhitalWP}. \\
$\bullet$ A nominal deep-drilling observing strategy over the full L/SMC galaxies will characterize stellar variability to $M_V<6.5$ on timescales from 15s to 3d \citep{SzkodyWP}. For this, special co-adds may be required, e.g., \textit{"to reach variability levels of 0.1 to 0.005 mag will require co-adds depending on the timescale of the particular variables"}. \citep{SzkodyWP}. \\
$\bullet$ A Twilight Short Exposure survey would enable bright stars, which have longer monitoring baselines due to historical observations, to be put on the same photometric system as the deeper LSST WFD main survey catalog. \\
$\bullet$ A short-exposure survey of M67, Chapter 10.4 of \citep{2017arXiv170804058L}, suggests using the stretch goal of 0.1 second exposures or, if that is not possible, \textit{"custom pixel masks to accurately perform photometry on stars as much as 6 magnitudes brighter than the saturation level"}. \\
$\bullet \bullet \bullet$ \textbf{Summary.} Many of these science goals can be achieved with standard visit images, but with alternative observing strategies such as limiting the number of filters or concentrating the depth build-up into shorter timeframes; the exception is short-exposure surveys to obtain bright stars. Similarly, most of these science goals can be met by reconfiguring the planned DM pipelines to detect sources, measure proper motion, and create deep stacks. The exception might be in processing exposures as short as 0.1 seconds, and attempting to measure flux of a saturated stars. See Section \ref{ssec:SPCS_GPVSEx} for more detailed DM processing case studies.

\noindent \textbf{Exoplanets} \\
$\bullet$ Transits. The nominal DDF plan described in \citep{2008arXiv0805.2366I} would allow for $1\%$ variability detection over hour-long timescales, which is suitable for detecting transits. A DDF field at Galactic latitude $30$ degrees would yield $10^6$ stars at $r<21$ that would have $\mathrm{SNR}>100$ in each single exposure of the sequence (as described in Section 3.1.2 of \citep{2008arXiv0805.2366I}). \citep{2013arXiv1304.3455G} describes how transits can be extract from a wider-area survey of the Galactic Plane.\\
$\bullet$ Microlensing.  Slower than a transit, \citep{2013arXiv1304.3455G} suggests that $\sim22$ mag imaging over the Galactic Plane region every 3-4 days (i.e., the WFD nominal cadence) can find microlensing candidates, which would then need follow-up with external facilities. However, this will require image differencing in crowded fields. \citep{2013arXiv1304.3455G} suggests that the Galactic Plane can yield a lot of science despite the fact that its eventual deep co-adds would be uselessly confusion limited, and therefore should not be skipped. Microlensing can also be done with DDF-style observing at low Galactic latitudes. \\
$\bullet \bullet \bullet$ \textbf{Summary.} All of these science goals appear possible with standard visit images. They might be possible with reconfigured DM pipelines, but this depends heavily on performance in crowded fields. See Section \ref{ssec:SPCS_GPVSEx} for a more detailed DM processing case study for Galactic Plane regions.

\noindent \textbf{Supernovae} \\
$\bullet$ The nominal DDF plan described in \citep{2008arXiv0805.2366I}, which builds nightly stacks with a limit of $r\sim26.5$ out of standard visit images, would extend the SN sample to $z\sim1.2$ and provide more densely sampled light curves for cosmological analyses. The optimal exposure time distribution might be 6, 5, 10, 10, 9, 10 in $ugrizy$. \citep{KesslerWP} \\
$\bullet$ High-cadence observations of DDF would be the only way to detect fast transients, particularly extragalactic novae, some tidal disruption events, optical counterparts to gamma-ray bursts, and peculiar SNe \citep{2014ApJ...794...23D}. \\
$\bullet$ Generating the best-possible individual SN light curves for cosmological analyses requires building special, deep-as-possible, SN-free host galaxy images and using them as a template. This will also be necessary for studying SNe that appear in the template image; i.e., that last $>1000$ days. These are mostly Type IIn, probably explosions of massive stars into dense circumstellar material, which are not used for cosmology but rather to study late-stage stellar evolution and mass loss. SN-free images will also be needed to measure correlated properties for cosmology and to do host-galaxy science. The latter, specifically the "characterization of ultra-faint SN host galaxies", is also mentioned in the Galaxies DDF WP \citep{FergusonWP}. \\
$\bullet$ \textit{MLG - I'm not sure this has been yet proposed, but short-exposure observations to include nearby SNe on the same photometric system could be useful?} \\
$\bullet \bullet \bullet$ \textbf{Summary.} All of these science goals appear possible with standard visit images (with the exception of an as-yet unproposed target-of-opportunity survey to obtain photometry for nearby, saturated SNe). Additionally, all are accessible with reconfigured DM pipelines to stack and difference the data. In particular, the Level 2 DRP codes to create "transient-free CoAdds" will be suitable for generating the SN-free templates for DDF, as they will do for the Main Survey images. See also Section \ref{ssec:SPCS_SNDDF} for a DM processing case study to find SNe in a DDF.

\noindent \textbf{Galaxies} \\
$\bullet$ Use the additional depth of a DDF to build a large collection of low-$\mu$ objects. \citep{FergusonWP} mentions "identification of nearby isolated low-redshift dwarf galaxies via surface-brightness fluctuations" and "characterization of low-surface-brightness extended features around both nearby and distant galaxies". \\
$\bullet$ The DDF stacks can also be used to characterize of high-$z$ clusters, although this ability might depend on deblending extended objects. \\
$\bullet$ The DDF observations, when combined with the WFD, allow for AGN monitoring on a variety of timescales in well-characterized galaxies \citep{FergusonWP,GawiserWP}. \\
$\bullet \bullet \bullet$ \textbf{Summary.} As with the SN science goals, these use standard visit images and reconfigured DM pipelines to make deep CoAdds and extract sources. In addition, it seems likely that Level 3 algorithms that are optimized to detect and characterize particular types of faint extended sources will be needed, but these are beyond the scope of DM.

@@

The document "Cadence Exploration and Improvement Plan" (by Zeljko, Lynne, Andy) proposes that the next call for DDF White Papers (proposed fields and cadences) be in December 2017, and the call for Mini-Survey White Papers be in October 2018. At the 2017 All Hands Meeting the Science Advisory Council seemed likely to recommend these be combined into a single call. The relevant Jira issue is \url{https://jira.lsstcorp.org/browse/LIT-325?filter=-3} and the Project's initial thoughts regarding this call for white papers are described by Beth Willman at the LSST SAC Meeting in October 2016\footnote{\url{https://project.lsst.org/groups/sac/sites/lsst.org.groups.sac/files/Willman_DDF.pdf}}. The plan for ingesting these white paper proposals is for them to be reviewed by the Science Advisory Council based on a set of criteria set by the Project Science Team, and the SAC will make recommendations to the LSST Director. The criteria to be met by the proposals may include specifications that they: satisfy minimum technical requirements (feasibility, overheads); maximize diverse scientific objectives (serve a wide community); and generate legacy datasets and add value to the products of other facilities. After the proposals have been reviewed (and perhaps culled or combined, as recommended by the SAC) it is projected that new OpSim runs will be produced in mid-2019 for further evaluation. While the initial white paper proposals will be separate from the existing Observing Strategy White Paper \citep{2017arXiv170804058L}, they could be incorporated into a comprehensive analysis of the options when these simulations are ready.

@@

A short-exposure survey of the bright stars of M67, described in Chapter 10.4 of the Observing Strategy White Paper \citep{2017arXiv170804058L}, suggests using the stretch goal of 0.1 second exposures or, if that is not possible, \textit{"custom pixel masks to accurately perform photometry on stars as much as 6 magnitudes brighter than the saturation level"}. This would certainly be considered a Level 3 algorithm.
